Trauma treatment delay

### 2023-10-25 manuscriptlongversion.Rmd

---
title: "Trauma treatment delay"
author: "Anton Wasielewski"
subtitle: A retrospective cohort study on the factors affecting the delays in trauma
  care
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
bibliography: bibliography.bib
editor_options:
  markdown:
    wrap: 72
---

@@

**Definition**

@@

Trauma systems are infrastructures that exist to provide and optimize
care for injured patients starting with injury recognition and triage,
transport to appropriate trauma center, inpatient care and outpatient
follow-up. Beyond the clinical side, trauma systems work with outreach,
education and advocacy, data collection through registries, research,
funding, and disaster preparedness and response. A comprehensive and
functioning trauma system requires strong leadership and engagement at
the trauma center, regional and national level [@connolly_2018]. This
system is crucial to provide care for trauma patients, both in reducing
morbidity and mortality in this patient group. Earlier studies from
Sweden have shown that treating severely injured patients at a trauma
center is associated with a 41% lower adjusted 30-day mortality rate
compared to being treated at a non trauma care center due to them being
more capable of treating these patients, with potential survival benefit
increasing with higher injury severity [@Candefjord_2020]. Other studies
have shown similar results, showing that treating injured patients at
trauma centers is associated with a 15-30% decrease in mortality
[@Moran_2018][@Celso_2006][@MacKenzie_2006][@Haas_2012]. During the last
three decades, the introduction of trauma systems has contributed to
lowering the incidence of preventable death . This is attributed to the
improvements in care for acute brain injuries and bleeding control. The
incidence of late death because of sepsis and multiple organ failure,
possibly a result of better and earlier resuscitation [@Asensio_2008].
Functioning trauma systems saves lives in the hospital, but their work
outside of the hospital is just as important. Data collection from
injured patients, such as mechanism of injury or mortality, are
essential for creating databases that can be used for research. In turn,
that research can be used for planning injury prevention programs that
target the most common injuries in the right ways, e.g. teen drivers,
children, specific occupations etc. These injury prevention programs can
be planned on trauma center, organizational or government level[@ACS].

@@

**Setting**

@@

**Study design**

@@

**Outcome**

@@

**Participants**

@@

**Variables**

@@

**Data sources/measurement**

@@

**Bias**

@@

**Study size**

@@

**Statistical methods**

>>>>

---
title: "Trauma treatment delay"
author: "Anton Wasielewski"
subtitle: A retrospective cohort study on the factors affecting the delays in trauma
  care
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
bibliography: bibliography.bib
csl: vancouver.csl
editor_options:
  markdown:
    wrap: 72
---

@@

## Definition

@@

Trauma systems are infrastructures that exist to provide and optimize
care for injured patients starting with injury recognition and triage,
transport to appropriate trauma center, inpatient care and outpatient
follow-up. Beyond the clinical side, trauma systems work with outreach,
education and advocacy, data collection through registries, research,
funding, and disaster preparedness and response. A comprehensive and
functioning trauma system requires strong leadership and engagement at
the trauma center, regional and national level [@connolly_2018]. This
system is crucial to provide care for trauma patients, both in reducing
morbidity and mortality in this patient group. Earlier studies from
Sweden have shown that treating severely injured patients at a trauma
center is associated with a 41% lower adjusted 30-day mortality rate
compared to being treated at a non trauma care center due to them being
more capable of treating these patients, with potential survival benefit
increasing with higher injury severity [@Candefjord_2020]. Other studies
have shown similar results, showing that treating injured patients at
trauma centers is associated with a 15-30% decrease in mortality
[@Moran_2018][@Celso_2006][@MacKenzie_2006][@Haas_2012]
<!-- KS: multiple citations should be like this: [@Moran_2018; @Celso_2006; @MacKenzie_2006; @Haas_2012]. Fix this where applicable -->.
During the last three decades, the introduction of trauma systems has
contributed to lowering the incidence of preventable death . This is
attributed to the improvements in care for acute brain injuries and
bleeding control. The incidence of late death because of sepsis and
multiple organ failure, possibly a result of better and earlier
resuscitation [@Asensio_2008]. Functioning trauma systems saves lives in
the hospital, but their work outside of the hospital is just as
important. Data collection from injured patients, such as mechanism of
injury or mortality, are essential for creating databases that can be
used for research. In turn, that research can be used for planning
injury prevention programs that target the most common injuries in the
right ways, e.g. teen drivers, children, specific occupations etc. These
injury prevention programs can be planned on trauma center,
organizational or government level[@ACS].

@@

## Setting

@@

## Study design

@@

## Outcome

@@

## Participants

@@

## Variables

@@

## Data sources/measurement

@@

## Bias

@@

## Study size

@@

## Statistical methods
